#Reproducible Research Peer Assessment 1
===========

This script loads activity.csv to variable, activity, then
performs analysis on steps variable based on different categories.

First, let's load activity.csv and check its structure.
```{r}
activity=read.csv(unz('activity.zip','activity.csv'),stringsAsFactor=F)
str(activity)
```

We have to convert date character to date type.
```{r}
activity$date=as.Date(activity$date,format="%Y-%m-%d")
```

Now compute total number of steps for each day and plot histogram.
```{r}
activity.dailyTotalSteps=tapply(activity$steps,activity$date,sum)
hist(activity.dailyTotalSteps)
```

The mean and median of activity.dailyTotalSteps.
```{r}
mean(activity.dailyTotalSteps,na.rm=T)
median(activity.dailyTotalSteps,na.rm=T)
```

Plot the number of steps per 5-minute, averaged across all days.
```{r}
steps.mean=tapply(activity$steps,activity$interval,mean,na.rm=T)
plot(steps.mean,type="l",xaxt="n",xlab="Hour of day",ylab="Average number of steps")
xtk=seq(0,288,12)
axis(1,at=xtk,labels=xtk/12)
```

Report which 5-minute interval, on average, has the maximum steps and the index of that max in steps.mean array.
```{r}
which.max(steps.mean)
```

How many missing values (NA) do we have in this activity?
```{r}
sum(is.na(activity$steps))
```

Replace NA with previous value in time series of the new data frame, imputedActivity.
```{r}
na.location=is.na(activity$steps)
imputedActivity=activity
if (na.location[1]) {
  lastSteps=0
  imputedActivity$steps[1]=0
} else {
  lastSteps=imputedActivity$steps[1]
}
na.location[1]
imputedActivity$steps[1]
for (i in seq(2,nrow(imputedActivity))) {
  if (na.location[i]) {
    imputedActivity$steps[i]=lastSteps
  } else {
    lastSteps=imputedActivity$steps[i]
  }
}
sum(is.na(imputedActivity$steps))
```

Now compute total number of steps for each day and plot histogram for imputedActivity.
```{r}
imputedActivity.dailyTotalSteps=tapply(imputedActivity$steps,imputedActivity$date,sum)
hist(imputedActivity.dailyTotalSteps)
```

The mean and median of activity.dailyTotalSteps.
```{r}
mean(imputedActivity.dailyTotalSteps,na.rm=T)
median(imputedActivity.dailyTotalSteps,na.rm=T)
```

We can clearly see mean and median values change after we impute.  Both of them become larger.

Let's compare activities between weekdays and weekends.
First, partition days into weekday and weekend.
```{r}
imputedActivity$day=weekdays(imputedActivity$date)
imputedActivity$day[imputedActivity$day=="Sunday" | imputedActivity$day=="Saturday"]="weekend"
imputedActivity$day[imputedActivity$day!="weekend"]="weekday"
imputedActivity$day=as.factor(imputedActivity$day)
```

Plot average number of steps for each 5-min interval, averaging over weekdays, and over weekends.
Note that we need to load lattice plotting system library.
```{r}
library(lattice)
weekdayActivity=subset(imputedActivity,imputedActivity$day=="weekday")
weekendActivity=subset(imputedActivity,imputedActivity$day=="weekend")
s1=tapply(weekdayActivity$steps,weekdayActivity$interval,mean)
s2=tapply(weekendActivity$steps,weekendActivity$interval,mean)
meanSteps=c(s1,s2)
interval=c(seq(0,1435,5),seq(0,1435,5))
day=c(rep("weekday",288),rep("weekend",288))
plotData=data.frame(meanSteps,interval,day)
plotData$day=as.factor(plotData$day)
xyplot(meanSteps ~ interval|day, data=plotData,type="l",ylab="Average number of steps in 5-min interval",xlab="index of 5-min intervals",layout=c(1,2))
```

We can clearly see the difference between these two plots.